---
title: "RRF Notebook"
output: html_notebook
---

## ------------------------------------------------- Import Libraries (REQUIRED) -------------------------------------------------

```{r}
library(readr)
library(plotrix)
```

## ----------------------------------------- Import the Dataset from Your Computer (REQUIRED) --------------------------------------

Change the location in the CSV file 
Ignore the "Missing column names" message if its appears

```{r}
RUSA_Recovery_Fund_Totals <- read_csv("/Users/fauzanamjad/Desktop/RUSA_Recovery_Fund_Totals.csv")
```

## ------------- Subset Dataset based on Class Year - Applying Students for Each Class ------------- 

For future use of this notebook, add subset functions that reflect all the class years on the application.

```{r}
class2021 <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "2021")
class2022 <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "2022")
class2020 <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "2020")
class2023 <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "2023")
class2024 <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "2024")
classGRAD <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Class Year:` == "GRAD")
```

## Subset Class Years by Whether they were Recommended - Recommended Students for Each Class Year

For future use of this notebook, change the subset functions/variables to reflect changes made to previous block of code.

```{r}
class2021Recommended <- subset(class2021, class2021$`Recommend/Denied` == "Recommend")
class2022Recommended <- subset(class2022, class2022$`Recommend/Denied` == "Recommend")
class2020Recommended <- subset(class2020, class2020$`Recommend/Denied` == "Recommend")
class2023Recommended <- subset(class2023, class2023$`Recommend/Denied` == "Recommend")
class2024Recommended <- subset(class2024, class2024$`Recommend/Denied` == "Recommend")
classGRADRecommended <- subset(classGRAD, classGRAD$`Recommend/Denied` == "Recommend")
```

## Number of Applying Students for Each Class Year VS Number of Recommended Students for Each Class Year (TABLE)

Adjust as needed with changes to the previous two blocks. 
nrow() is the function that tells how many elements (students) are in that specific dataset.
We store these values in a columns, than combine these columns into a dataframe. 

```{r}
Class <- c("2021", "2022", "2020", "2023","2024", "Grad")
Frequency_of_Class <- c(nrow(class2021), nrow(class2022), nrow(class2020), nrow(class2023), nrow(class2024), nrow(classGRAD))
Recommended_Students <- c(nrow(class2021Recommended), nrow(class2022Recommended), nrow(class2020Recommended), nrow(class2023Recommended), nrow(class2024Recommended), nrow(classGRADRecommended))
class_breakdown <- data.frame(Class,Frequency_of_Class ,Recommended_Students)
class_breakdown
```

## Export 

This code exports a csv file of the class year breakdown so it can be used further in excel.
Change the Destination to a folder than you want to store the CSV file.

```{r}
write.csv(class_breakdown, "/Users/fauzanamjad/Desktop/ClassBreakdown.csv", row.names = FALSE)
```


## -----------------------  Subset data based on school - Requesting Students by School -----------------------

Add more schools as need if reflected in the RRF dataset or application. 

```{r}
classBloustein <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "Bloustein")
classEMSOP <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "EMSOP")
classMGSA <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "MGSA")
classNA <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "N/A")
classNursing <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "Nursing")
classRBS <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "RBS")
classSAS <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "SAS")
classSOE <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "SOE")
classSEBS <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`School:` == "SEBS")
```

## Subset School by Whether or not they were Recommended - Recommended Students by School

If you added schools in the previous block, remember to make the respective change here as well. 

```{r}
classBloustein_Rec <- subset(classBloustein, classBloustein$`Recommend/Denied` == "Recommend")
classEMSOP_Rec <- subset(classEMSOP, classEMSOP$`Recommend/Denied` == "Recommend")
classMGSA_Rec <- subset(classMGSA, classMGSA$`Recommend/Denied` == "Recommend")
classNA_Rec <- subset(classNA, classNA$`Recommend/Denied` == "Recommend")
classNursing_Rec <- subset(classNursing, classNursing$`Recommend/Denied` == "Recommend")
classRBS_Rec <- subset(classRBS, classRBS$`Recommend/Denied` == "Recommend")
classSOE_Rec <- subset(classSOE, classSOE$`Recommend/Denied` == "Recommend")
classSEBS_Rec <- subset(classSEBS, classSEBS$`Recommend/Denied` == "Recommend")
classSAS_Rec <-subset(classSAS, classSAS$`Recommend/Denied` == "Recommend")
```

## Number of Applying Students by School VS Number of Recommended Students by School

Adjust as needed with changes to the previous two blocks. 
nrow() is the function that tells how many elements (students) are in that specific dataset.
We store these values in a columns, than combine these columns into a dataframe. 

```{r}
School <- c("Bloustein", "EMSOP", "MGSA", "N/A", "Nursing", "RBS", "SAS", "SOE","SEBS")
School
Frequency <- c(nrow(classBloustein), nrow(classEMSOP),nrow(classMGSA),nrow(classNA),nrow(classNursing),nrow(classRBS),nrow(classSAS),nrow(classSOE),nrow(classSEBS))
Frequency
Recommended_Students1 <- c(nrow(classBloustein_Rec), nrow(classEMSOP_Rec),nrow(classMGSA_Rec),nrow(classNA_Rec),nrow(classNursing_Rec),nrow(classRBS_Rec),nrow(classSAS_Rec),nrow(classSOE_Rec),nrow(classSEBS_Rec))
breakdown_school <- data.frame(School,Frequency,Recommended_Students1)
breakdown_school
```

## Piechart of School Breakdown of Requesting Students

Not the best pie chart you can make, but if you need something quick this should be suffice. 

```{r}
pct <- round(Frequency/sum(Frequency)*100)
School <- paste(School, pct) # add percents to labels
School <- paste(School,"%",sep="") # ad % to labels
pie(Frequency, labels = School, main="Breakdown by School: All Applications", col=rainbow(length(School)))
```

## Export School Breakdown - Requesting & Recommended

This code exports a csv file of the school breakdown so it can be used further in excel.
Change the Destination to a folder than you want to store the CSV file.

```{r}
write.csv(breakdown_school, "/Users/fauzanamjad/Desktop/SchoolBreakdown.csv", row.names = FALSE)
```

##  -----------------------  Subset by Whether or not they're on DACA - Requesting Students on or not on DACA -----------------------  

```{r}
yesDACA <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`DACA (Y/N)` == "Y")
noDACA <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`DACA (Y/N)` == "N")
unDACA <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`DACA (Y/N)` == "N/A")
```

## Subset DACA subsets by recommendation - Applying Students on or not on DACA

```{r}
yesDACA_rec <- subset(yesDACA, yesDACA$`Recommend/Denied` == "Recommend")
noDACA_rec <- subset(noDACA, noDACA$`Recommend/Denied` == "Recommend")
unDACA_rec <- subset(unDACA , unDACA$`Recommend/Denied` == "Recommend")
```

## DACA VS Not on DACA Students (Requesting and Recommended Data)

```{r}
DACA_Label <- c("YES", "NO", "N/A")
DACA_req_freq <- c( nrow(yesDACA) , nrow(noDACA) , nrow(unDACA) ) ## Requesting Students 
DACA_recom_frq <- c( nrow(yesDACA_rec) , nrow(noDACA_rec) , nrow(unDACA_rec) ) ## Recommended Students
DACA_breakdown <- data.frame(DACA_Label,DACA_req_freq,DACA_recom_frq) 
DACA_breakdown
```

## Piechart of DACA Breakdown

```{r}
pct <- round(DACA_req_freq/sum(DACA_req_freq)*100)
DACA_Label <- paste(DACA_Label, pct) # add percents to labels
DACA_Label <- paste(DACA_Label,"%",sep="") # ad % to labels
pie(DACA_req_freq, labels = DACA_Label, main="Breakdown by DACA: All Applications", col=rainbow(length(DACA_Label)))
```

## Export DACA Breakdown

```{r}
write.csv(DACA_breakdown, "/Users/fauzanamjad/Desktop/DACA_Breakdown.csv", row.names = FALSE)
```

##  -----------------  Subset by Whether or not they're International - Requesting Students on or not on International -------------------  

```{r}
yesInternational <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`International (Y/N)` == "Y")
noInternational<- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`International (Y/N)` == "N")
unInternational <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`International (Y/N)` == "N/A")
```

## Subset International subsets by recommendation - Applying Students and whether they're International

```{r}
yesInternational_rec <- subset(yesInternational, yesInternational$`Recommend/Denied` == "Recommend")
noInternational_rec <- subset(noInternational, noInternational$`Recommend/Denied` == "Recommend")
unInternational_rec <- subset(unInternational , unInternational$`Recommend/Denied` == "Recommend")
```

## International VS Non International Students (Requesting and Recommended Data)

```{r}
International_Label <- c("YES", "NO", "N/A")
International_req_freq <- c( nrow(yesInternational) , nrow(noInternational) , nrow(unInternational) ) ## Requesting Students 
International_recom_frq <- c( nrow(yesInternational_rec) , nrow(noInternational_rec) , nrow(unInternational_rec) ) ## Recommended Students
International_breakdown <- data.frame(International_Label,International_req_freq,International_recom_frq) 
International_breakdown
```

## Piechart of International Students (or nah) Breakdown

```{r}
pct <- round(International_req_freq/sum(International_req_freq)*100)
International_Label <- paste(International_Label, pct) # add percents to labels
International_Label <- paste(International_Label,"%",sep="") # ad % to labels
pie(International_req_freq, labels = International_Label, main="Breakdown by International Status: All Applications", col=rainbow(length(International_Label)))
```

## Export International Student Breakdown

```{r}
write.csv(International_breakdown, "/Users/fauzanamjad/Desktop/International_Breakdown.csv", row.names = FALSE)
```

##  -----------------------  Subset by Whether or not they're Food Insecure - Requesting Students by Food Insecurity -----------------------  

```{r}
yesFI <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Insecure (Y/n)` == "Y")
noFI <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Insecure (Y/n)` == "N")
unFI <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Insecure (Y/n)` == "N/A")
```

## Subset Food Insecure subsets by recommendation - Applying Students by Food Insecurity

```{r}
yesFI_rec <- subset(yesFI, yesFI$`Recommend/Denied` == "Recommend")
noFI_rec <- subset(noFI, noFI$`Recommend/Denied` == "Recommend")
unFI_rec <- subset(unFI , unFI$`Recommend/Denied` == "Recommend")
```

## Food Insecure VS Not Food Insecure Students (Requesting and Recommended Data)

```{r}
FI_Label <- c("YES", "NO", "N/A")
FI_req_freq <- c( nrow(yesFI) , nrow(noFI) , nrow(unFI) ) ## Requesting Students 
FI_recom_frq <- c( nrow(yesFI_rec) , nrow(noFI_rec) , nrow(unFI_rec) ) ## Recommended Students
FI_breakdown <- data.frame(FI_Label,FI_req_freq,FI_recom_frq) 
FI_breakdown
```

## Piechart of Food Insecurity Breakdown

```{r}
pct <- round(FI_req_freq/sum(FI_req_freq)*100)
FI_Label <- paste(FI_Label, pct) # add percents to labels
FI_Label <- paste(FI_Label,"%",sep="") # ad % to labels
pie(FI_req_freq, labels = FI_Label, main="Breakdown by Food Insecurity: All Applications", col=rainbow(length(FI_Label)))
```

## Export Food Insecurity Breakdown

```{r}
write.csv(FI_breakdown, "/Users/fauzanamjad/Desktop/FoodInsecurity_Breakdown.csv", row.names = FALSE)
```

## --------------------------------- REQUESTED FOOD MONEY SUBSET --------------------------------- 

```{r}
requestedFoodMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Amount:` > 0.00)
not_requestedFoodMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Amount:` == 0.00)
na_requestedFoodMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Food Amount:` == "N/A")
```

## Subsetting by recommended people (students who requested money for food)

```{r}
requestedFoodMoney_rec <- subset(requestedFoodMoney, requestedFoodMoney$`Recommend/Denied` == "Recommend")
not_requestedFoodMoney_rec <- subset(not_requestedFoodMoney, not_requestedFoodMoney$`Recommend/Denied` == "Recommend")
na_requestedFoodMoney_rec <- subset(na_requestedFoodMoney, na_requestedFoodMoney$`Recommend/Denied` == "Recommend")
```

## Table view of Applying VS Recommended Students by whether or not they requested money for food

```{r}
Requested_Food_Label <- c("YES", "NO", "N/A")
Requested_Food_req_freq <- c( nrow(requestedFoodMoney) , nrow(not_requestedFoodMoney) , nrow(na_requestedFoodMoney) ) ## Requesting Students 
Requested_Food_recom_frq <- c( nrow(requestedFoodMoney_rec) , nrow(not_requestedFoodMoney_rec) , nrow(na_requestedFoodMoney_rec) ) ## Recommended Students
Requested_Food_breakdown <- data.frame(Requested_Food_Label,Requested_Food_req_freq,Requested_Food_recom_frq) 
Requested_Food_breakdown
```

## Pie chart for Food Requesting Breakdown of Applying Students

```{r}
pct <- round(Requested_Food_req_freq/sum(Requested_Food_req_freq)*100)
Requested_Food_Label <- paste(Requested_Food_Label, pct) # add percents to labels
Requested_Food_Label <- paste(Requested_Food_Label,"%",sep="") # ad % to labels
pie(Requested_Food_req_freq, labels = Requested_Food_Label, main="Breakdown by Food Money Requests: All Applications", col=rainbow(length(Requested_Food_Label)))
```

## Export 

```{r}
write.csv(Requested_Food_breakdown, "/Users/fauzanamjad/Desktop/RequestedFood_Breakdown.csv", row.names = FALSE)
```

## --------------------------------- REQUESTED RENT MONEY SUBSET --------------------------------- 

```{r}
requestedRentMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Rent Amount:` > 0.00)
not_requestedRentMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Rent Amount:` == 0.00)
na_requestedRentMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Rent Amount:` == "N/A")
```

## Subsetting by recommended people (students who requested money for rent)

```{r}
requestedRentMoney_rec <- subset(requestedRentMoney, requestedRentMoney$`Recommend/Denied` == "Recommend")
not_requestedRentMoney_rec <- subset(not_requestedRentMoney, not_requestedRentMoney$`Recommend/Denied` == "Recommend")
na_requestedRentMoney_rec <- subset(na_requestedRentMoney, na_requestedRentMoney$`Recommend/Denied` == "Recommend")
```

## Table view of Applying VS Recommended Students by whether or not they requested money for rent

```{r}
Requested_Rent_Label <- c("YES", "NO", "N/A")
Requested_Rent_req_freq <- c( nrow(requestedRentMoney) , nrow(not_requestedRentMoney) , nrow(na_requestedRentMoney) ) ## Requesting Students 
Requested_Rent_recom_frq <- c( nrow(requestedRentMoney_rec) , nrow(not_requestedRentMoney_rec) , nrow(na_requestedRentMoney_rec) ) ## Recommended Students
Requested_Rent_breakdown <- data.frame(Requested_Rent_Label,Requested_Rent_req_freq,Requested_Rent_recom_frq) 
Requested_Rent_breakdown
```

## Pie chart for Rent Requesting Breakdown of Applying Students

```{r}
pct <- round(Requested_Rent_req_freq/sum(Requested_Rent_req_freq)*100)
Requested_Rent_Label <- paste(Requested_Rent_Label, pct) # add percents to labels
Requested_Rent_Label <- paste(Requested_Rent_Label,"%",sep="") # ad % to labels
pie(Requested_Rent_req_freq, labels = Requested_Rent_Label, main="Breakdown by Rent Money Requests: All Applications", col=rainbow(length(Requested_Rent_Label)))
```

## Export 

```{r}
write.csv(Requested_Rent_breakdown, "/Users/fauzanamjad/Desktop/RequestedRent_Breakdown.csv", row.names = FALSE)
```

## --------------------------------- REQUESTED OTHER MONEY SUBSET --------------------------------- 

```{r}
requestedOtherMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Other:` > 0.00)
not_requestedOtherMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Other:` == 0.00)
na_requestedOtherMoney <- subset(RUSA_Recovery_Fund_Totals, RUSA_Recovery_Fund_Totals$`Other:` == "N/A")
```

## Subsetting by recommended people (students who requested money for other)

```{r}
requestedOtherMoney_rec <- subset(requestedOtherMoney, requestedOtherMoney$`Recommend/Denied` == "Recommend")
not_requestedOtherMoney_rec <- subset(not_requestedOtherMoney, not_requestedOtherMoney$`Recommend/Denied` == "Recommend")
na_requestedOtherMoney_rec <- subset(na_requestedOtherMoney, na_requestedOtherMoney$`Recommend/Denied` == "Recommend")
```

## Table view of Applying VS Recommended Students by whether or not they requested money for other

```{r}
Requested_Other_Label <- c("YES", "NO", "N/A")
Requested_Other_req_freq <- c( nrow(requestedOtherMoney) , nrow(not_requestedOtherMoney) , nrow(na_requestedOtherMoney) ) ## Requesting Students 
Requested_Other_recom_frq <- c( nrow(requestedOtherMoney_rec) , nrow(not_requestedOtherMoney_rec) , nrow(na_requestedOtherMoney_rec) ) ## Recommended Students
Requested_Other_breakdown <- data.frame(Requested_Other_Label,Requested_Other_req_freq,Requested_Other_recom_frq) 
Requested_Other_breakdown
```

## Pie chart for Other Requesting Breakdown of Applying Students

```{r}
pct <- round(Requested_Other_req_freq/sum(Requested_Other_req_freq)*100)
Requested_Other_Label <- paste(Requested_Other_Label, pct) # add percents to labels
Requested_Other_Label <- paste(Requested_Other_Label,"%",sep="") # ad % to labels
pie(Requested_Other_req_freq, labels = Requested_Other_Label, main="Breakdown by Money Requests for Other Needs: All Applications", col=rainbow(length(Requested_Other_Label)))
```

## Export 

```{r}
write.csv(Requested_Other_breakdown, "/Users/fauzanamjad/Desktop/RequestedOther_Breakdown.csv", row.names = FALSE)
```

